### 2023-07-17 paper/sec-dis-pseudodata.tex
Update sec-dis-pseudodata.tex

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{eq:fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
We take the uncertainty associated
to a given systematic source, say $\delta E_\ell$,
to be the absolute difference between the mean number of events in this
bin evaluated for the smeared samples samples and the
central prediction for the number of events
$ N_{\rm ev}^{(i)}$ as calculated from Eq.~(\ref{MCintegration}):
\be
\delta_{\rm sys}^{(E_\ell)} N^{(i)} = \Big| \la {N}^{(i)}\ra_{E_\ell-{\rm smeared}} -N_{\rm ev}^{(i)}\Big| \, .
\ee
Systematic errors are treated as uncorrelated among them and hence
by producing samples where only one source of error is varied at a time
we can determine the systematic errors, Eq.~(\ref{eq:event_yields_systematic_error}), in each bin.
%
This approach has the benefit that rescaling individual sources of systematic
uncertainties, say to assess the impact of improved detector performance,
are straightforward.

Fig.~\ref{fig:percentage_uncertainties_overview}
displays the systematic uncertainties associated
to $E_\ell$, $\theta_\ell$, and $E_h$.
in the  measurements
of the double-differential
neutrino scattering cross-section at FASER$\nu$2.
%
The magnitude of each systematic error is plotted as a function
of the average momentum fraction per bin $\la x\ra$
in three different bins of $Q^2$.
%
We indicate separately the results for neutrino and antineutrino projectiles as well as
those associated to inclusive and to charm production measurements.
%
Systematic uncertainties associated to the hadronic final state energy $E_h$ appear to be the largest.
%
The uncertainties associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$ range
between a few percent up to 20\% in the nominal baseline scenario,
and in the case of the latter they are the smallest in the large-$x$ region.
%
The same procedure is adopted for all the experiments listed in Table~\ref{tab:FPF_experiments}.

%-----------------------------------------------------------------------
\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{plots/percentage_errors.pdf}
  \caption{\small Estimated systematic uncertainties for the  measurements
    of the double-differential
    neutrino scattering cross-section at FASER$\nu$2.
    %
    We consider the systematic errors
    associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$
    and to the hadronic energy $E_h$.
    %
    The size of each source of systematic error is plotted as a function
    of the average momentum fraction per bin $\la x\ra$
    in three different bins of $Q^2$.
    %
    We indicate separately the results for neutrino and antineutrino projectiles as well as
    those associated to inclusive and to charm production measurements.
  }
  \label{fig:percentage_uncertainties_overview}
\end{figure}
%-----------------------------------------------------------------------

The end result of the procedure is the estimate of statistical and systematic uncertainties
for each bin of the measurement, from which a experimental covariance matrix can be constructed as
\be
   {\rm cov}_{ij} = \delta_{ij} \lp \delta_{\rm stat}  N_{\rm ev}^{(i)}\rp^2
   + \lp \delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(i)} \rp \lp \delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(j)} \rp 
   + \lp \delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(i)} \rp \lp\delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(j)}\rp 
   + \lp \delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(i)} \rp
   \lp \delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(j)} \rp
   \, ,\qquad
 \nonumber
 \ee
 for $i,j=1,\ldots,N_{\rm bin}$, and the same for the associated correlation
 matrix of the measurement
 \be
 \rho_{ij} =  \frac{{\rm cov}_{ij}}{\sqrt{ {\rm cov}_{ii} }\sqrt{ {\rm cov}_{jj} } } \, . 
 \ee
 The relative covariance matrix, $ {\rm cov}_{ij}/( N_{\rm ev}^{(i)}N_{\rm ev}^{(j)})$, is
 independent of the considered observable and would also apply
 for the double-differential cross-sections Eqns.~(\ref{eq:neutrino_DIS_xsec_FL}) and~(\ref{eq:antineutrino_DIS_xsec_FL}) which are related to the event yields by a constant factor.

 One should emphasize that in a real experiment the actual covariance matrix will be
 composed by a much larger number of uncertainty sources, with typical
  HERA and LHC precision measurements characterised by up to hundreds
 of different sources of systematic error.
 %
 In particular, the assumption that a single source of systematic error, say $\delta E_\ell$,
 is fully correlated among all the bins in $(x,Q^2)$ is likely not to be realistic.
 %
 For this reason, here we present impact results both for the nominal correlation model,
 for the case where systematic and statistical uncertainties are added in quadrature,
 and intermediate scenarios.

 Fig.~\ref{fig:error_plot_FASERv2_14} (central and right panels) display
 the integrated event yields for FASER$\nu$2 as a function of $x$ for only
 systematic errors and for the sum in quadrature of statistical and systematic errors.
 %
 For most of the bins, for the baseline performance assumptions the total systematic
 uncertainty is at the $10\%$ and dominates over the statistical uncertainties.

 \subsection{Pseudo-data generation and impact on PDFs}

 In order to generate pseudo-data for double-differential
 neutrino scattering cross-sections at the LHC, we follow the procedure
 used for the HL-LHC PDF projections of~\cite{AbdulKhalek:2018rok} which was
 also adopted in~\cite{Ethier:2021ydt} and~\cite{Greljo:2021kvv} for SMEFT impact projections
 of vector-boson scattering and high-mass Drell-Yan data at the HL-LHC, respectively.
 %
 The starting point are the predictions for the differential neutrino scattering
 cross-section
 \be
 \label{eq:theory_dis_projections}
 \mathcal{O}_i^{{\rm (th)}} \equiv \frac{d^2\sigma^{\nu N}(x_i,Q^2_i,y_i)}{dxdy} \, ,\quad
 i=1,\ldots,N_{\rm bin} \, ,
 \ee
 with $(x_i,Q^2_i,y_i)$ being the corresponding bin centers.
 %
 The observables $\mathcal{O}_i $ are evaluated using {\sc\small YADISM} and {\sc\small PineAPPL}
 as described in Sect.~\ref{sec:settings}.
 %
 An important point here is the choice of PDF set entering the evaluation of
 the observable in Eq.~(\ref{eq:theory_dis_projections}): it should be consistent
 with that of the fitting framework used to assess their impact on the (n)PDFs.
 %
 For instance, when using the {\sc\small xFitter} profiling of PDF4LHC21, one needs
 to generate LHC neutrino pseudo-data also using PDF4LHC21 as input.
 %
 In order words, the generated pseudo-data should be fully consistent with the prior PDF
 set used as baseline, else one is introducing artificially inconsistencies which
 compromise the validity of the projection studies.

>>>>

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{eq:fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
We take the uncertainty associated
to a given systematic source, say $\delta E_\ell$,
to be the absolute difference between the mean number of events in this
bin evaluated for the smeared samples samples and the
central prediction for the number of events
$ N_{\rm ev}^{(i)}$ as calculated from Eq.~(\ref{MCintegration}):
\be
\delta_{\rm sys}^{(E_\ell)} N^{(i)} = \Big| \la {N}^{(i)}\ra_{E_\ell-{\rm smeared}} -N_{\rm ev}^{(i)}\Big| \, .
\ee
Systematic errors are treated as uncorrelated among them and hence
by producing samples where only one source of error is varied at a time
we can determine the systematic errors, Eq.~(\ref{eq:event_yields_systematic_error}), in each bin.
%
This approach has the benefit that rescaling individual sources of systematic
uncertainties, say to assess the impact of improved detector performance,
is straightforward.

Fig.~\ref{fig:percentage_uncertainties_overview}
displays the systematic uncertainties associated
to $E_\ell$, $\theta_\ell$, and $E_h$ variations
for the  measurements
of the double-differential
neutrino scattering cross-section at FASER$\nu$2.
%
The magnitude of each systematic error is plotted as a function
of the average momentum fraction per bin $\la x\ra$
in three different bins of $Q^2$.
%
We indicate separately the results for neutrino and antineutrino projectiles as well as
those associated to inclusive and to charm production measurements.
%
Systematic uncertainties associated to the hadronic final state energy $E_h$ appear to be the largest.
%
The uncertainties associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$ range
between a few percent up to 20\% in the nominal baseline scenario,
and in the case of the latter they are the smallest in the large-$x$ region.
%
Fig.~\ref{fig:error_plot_FASERv2_14} (central and right panels) display
 the integrated event yields for FASER$\nu$2 as a function of $x$ for only
 systematic errors and for the sum in quadrature of statistical and systematic errors.
 %
 For most of the bins, for the baseline performance assumptions the total systematic
 uncertainty is at the $10\%$ and dominates over the statistical uncertainties.

%-----------------------------------------------------------------------
\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{plots/percentage_errors.pdf}
  \caption{\small Estimated systematic uncertainties for the  measurements
    of the double-differential
    neutrino scattering cross-section at FASER$\nu$2.
    %
    We consider the systematic errors
    associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$
    and to the hadronic energy $E_h$.
    %
    The size of each source of systematic error is plotted as a function
    of the average momentum fraction per bin $\la x\ra$
    in three different bins of $Q^2$.
    %
    We indicate separately the results for neutrino and antineutrino projectiles as well as
    those associated to inclusive and to charm production measurements.
  }
  \label{fig:percentage_uncertainties_overview}
\end{figure}
%-----------------------------------------------------------------------

The end result of the procedure is the estimate of statistical and systematic uncertainties
for each bin of the measurement, from which a experimental covariance matrix can be constructed as
\be
\label{eq:covmat_definition}
   {\rm cov}_{ij} = \delta_{ij} \lp \delta^{\rm (stat)}  N_{\rm ev}^{(i)}\rp^2
   + \sum_{k=1}^{n_{\rm sys}}\lp \delta_{\rm sys}^{(k)} N_{\rm ev}^{(i)} \rp \lp \delta_{\rm sys}^{(k)} N_{\rm ev}^{(j)} \rp
   \, ,\qquad i,j=1,\ldots,N_{\rm bin} \, ,
 \ee
  and the same for the associated correlation
 matrix of the measurement
 \be
 \rho_{ij} =  \frac{{\rm cov}_{ij}}{\sqrt{ {\rm cov}_{ii} }\sqrt{ {\rm cov}_{jj} } } \, . 
 \ee
 The relative covariance matrix, $ {\rm cov}_{ij}/( N_{\rm ev}^{(i)}N_{\rm ev}^{(j)})$, is
 independent of the considered observable and would also apply
 for the double-differential cross-sections Eqns.~(\ref{eq:neutrino_DIS_xsec_FL}) and~(\ref{eq:antineutrino_DIS_xsec_FL}) which are related to the event yields by a constant prefactor.
 %

 Eq.~(\ref{eq:covmat_definition}) assumes that each source of systematic
 uncertainty is 100\% correlated across all the bins of the measurement.
 %
 In the real experiment, the actual covariance matrix will be
 composed by a large number of uncertainty sources, with typical
  HERA and LHC precision measurements characterised by up to hundreds
 of different sources of systematic error.
 %
 In particular, the assumption that a single source of systematic error, say $\delta E_\ell$,
 is fully correlated among all the bins in $(x,Q^2)$ is unlikely to be accurate.
 %
 For this reason, here we neglect bin-by-bin correlations
 and add in quadrature statistical and systematic errors
 \be
\label{eq:covmat_definition_v2}
 {\rm cov}_{ij} = \delta_{ij} \lp \delta^{\rm (stat)}  N_{\rm ev}^{(i)}\rp^2
+ \delta_{ij}\sum_{k=1}^{n_{\rm sys}} \lp f_{\rm red}^{(k)}\rp^2\lp \delta_{\rm sys}^{(k)} N_{\rm ev}^{(i)} \rp^2
\, ,\qquad i,j=1,\ldots,N_{\rm bin} \, .
\ee
In Eq.~(\ref{eq:covmat_definition_v2}) we have introduced a parameter $f_{\rm red}^{(k)}\le 1$
that gauges the impact of a possible reduction of the $k$-th systematic error
as compared to the default experiment performance.
%
In Sect.~\ref{sec:protonPDFs} we present results for different scenarios
for the systematic uncertainty reduction $f_{\rm red}^{(k)}$.

 \subsection{Pseudo-data generation and inclusion on PDF fits}

 In order to generate pseudo-data for double-differential
 LHC neutrino scattering cross-sections, we follow the procedure
 used for the HL-LHC PDF projections of~\cite{AbdulKhalek:2018rok} which was
 also adopted in~\cite{Ethier:2021ydt} and~\cite{Greljo:2021kvv} for SMEFT impact projections
 of vector-boson scattering and high-mass Drell-Yan data at the HL-LHC, respectively.
 %
 The starting point are the predictions for the differential neutrino scattering
 cross-section
 \be
 \label{eq:theory_dis_projections}
 \mathcal{O}_i^{{\rm (th)}} \equiv \frac{d^2\sigma^{\nu N}(x_i,Q^2_i,y_i)}{dxdy} \, ,\quad
 i=1,\ldots,N_{\rm bin} \, ,
 \ee
 with $(x_i,Q^2_i,y_i)$ being the corresponding bin centers.
 %
 The observables $\mathcal{O}_i $ are evaluated using {\sc\small YADISM} and {\sc\small PineAPPL}
 as described in Sect.~\ref{sec:settings}.
 %
 An important point here is the choice of PDF set entering the evaluation of
 the observable in Eq.~(\ref{eq:theory_dis_projections}): it should be consistent
 with that of the fitting framework used to assess their impact on the (n)PDFs.
 %
 For instance, when using the {\sc\small xFitter} profiling of PDF4LHC21, one needs
 to generate LHC neutrino pseudo-data also using PDF4LHC21 as input.
 %
 In order words, the generated pseudo-data should be fully consistent with the prior PDF
 set used as baseline, else one is introducing artificially inconsistencies which
 compromise the validity of the projection studies.
